Tail-indices and scale parameters of returns series

### 2017-04-05 papers/FX/draft.tex
Some more revision of the draft.

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
%\usepackage[bookmarks=true]{hyperref}
%\usepackage{bookmark}
\usepackage{graphicx}
\usepackage[bookmarks=true]{ hyperref}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

In Section~\ref{sec:4} we summarize the discussion of the previous sections.

@@

For convenience, we denote the scale parameter by $A$. It is the
constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
Hill \cite{hill1975simple} put forward an estimator of $A$ using his
tail index  estimator:
\[
\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
\]
where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
is the Hill tail index estimator.
Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in figure \ref{fig:Energy_scale},
\ref{fig:Consumer_Staples_scale} and
\ref{fig:Information_Technology_scale}, along with estimates obtained
using the hybrid method described in section
\ref{sec:hybrid_estimation}.

@@

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
increasing and concave. Before proceeding any further, it is worth
stating the following simple observation:
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
  $\E h(X) < \infty$ and
  \begin{eqnarray*}
    {\partial \over \partial \theta}\left[
      F(b) - F(a)
    \right] = 0 \\
    \left| \pd{f(x, \theta)}{\theta} \right| < \infty
    \quad \forall x \in (a, b)
  \end{eqnarray*}
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
    \[
    \pd{\E h(X)}{\theta} > 0
    \]
  \end{enumerate}
\end{lemma}
\begin{remark}
  \label{remark:I}
  Two other cases follow trivially from lemma \ref{lemma:I}:
  \begin{enumerate}
  \item If $h(\cdot)$ is increasing and $\pd{f}{\theta}$ satisfies the
    same conditions of the 1st case of lemma \ref{lemma:I},
    $\pd{\E h(X)}{\theta} < 0$. This immediately follows from applying
    1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
  \item By the same argument, if $h(\cdot)$ is decreasing and
    $\pd{f}{\theta}$ satisfies the same conditions of the 2nd case of
    lemma \ref{lemma:I}, $\pd{\E h(X)}{\theta} < 0$.
  \end{enumerate}
\end{remark}
Now for simplicity we assume the investor initially has one unit of
wealth. Let $r > 0$ be the return on the risk free state bond and $X$
be the return on equity. Portfolio weights are $\left(1-\phi\right)$
and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  C(X) = (1 - \phi) e^r + \phi e^X
\end{equation}
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q = q(r, \phi) &:=& \ln \left( {
      \delta v - (1 - \phi) e^r
      \over
      \phi
    } \right) = \ln\left(
    e^r + {\delta v - e^r \over \phi}
  \right)
\end{eqnarray*}
Note $C < \delta v$ implies $X < q$. Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher return from the risky asset than he is guarranteed from the
riskless bond. In our notation, this means $\delta v > e^r$, which
implies $q(r, \phi) > r > 0$. The right-hand-side of
\eqref{eq:functional} is a functional of $\phi$ and of the distribution
function of $X$ via \eqref{eq:xxie1}. For convenience we denote it
$\mathcal U(\cdot, \cdot)$, i.e.
\begin{equation}
  \label{eq:U_functional}
  \mathcal U(F, \phi)
  = 
  \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\end{equation}
In the rest of this section, we use $F(\cdot)$ to denote the
distribution function of $X$ and use $f(\cdot)$ to denote its density
function. Let
\begin{equation}
  \label{eq:G_functional}
  G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)  
\end{equation}

@@

\subsection{Hybrid fitting of the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:X_distr}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:X_distr}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$. Regularity
conditions on the maximum likelihood estimator are easily checked for
the distribution \eqref{eq:X_distr}.
The density function of $X$ conditional on $X < 0$ is
\[
f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
\]
Clearly, the log-likelihood function of the $K$ parameter is
\[
l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
\]
where, as is just mentioned, we pre-estimate $\alpha$ with Hill
estimator and regard it as known at this step. With $n$ observations
$X_1, \dots, X_n$, the log-likelihood function of $K$ is
\[
L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
\]
To obtain the maximum likelihood estimation, we need to solve the
equation $\pd{L}{K} = 0$, which expands to
\[
{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
\]
This equation has to be solved numerically.
The Fisher information of $K$ follows as
\begin{eqnarray*}
  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
  &=& \alpha K^{\alpha - 2}
  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
  &=& {\alpha \over (2 + \alpha) K^2}
\end{eqnarray*}
So the maximum likelihood estimator, call it $\hat K$, is
asymptotically normal:
\[
\sqrt n (\hat K - K) \to N \left(
  0, \left( 1 + {2 \over \alpha} \right) K^2
\right)
\]

@@

\begin{figure}[htb!]
  \centering
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Energy_p.pdf}
  %                   \caption{``Energy''}
  %                   \label{fig:Energy_p}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Energy_alpha_K_ci.pdf}
  %                   \caption{``Energy''}
  %                   \label{fig:Energy_alpha_K_ci}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Consumer_Staples_p.pdf}
  %                   \caption{``Consumer Staples''}
  %                   \label{fig:Consumer_Staples_p}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Consumer_Staples_alpha_K_ci.pdf}
  %                   \caption{``Consumer Staples''}
  %                   \label{fig:Consumer_Staples_alpha_K_ci}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Information_Technology_p.pdf}
  %                   \caption{``Information Technology''}
  %                   \label{fig:Information_Technology_p}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.45\linewidth}
  %   \includegraphics[width=\textwidth]
  %                   {Information_Technology_alpha_K_ci.pdf}
  %                   \caption{``Information Technology''}
  %                   \label{fig:Information_Technology_alpha_K_ci}
  % \end{subfigure}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_p.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
  \label{fig:sectors_parameters}
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
Comparing this expression with \eqref{eq:lower_tail_expansion}, one
can immediately identity $p K^\alpha$ as the scale parameter. Figure
\ref{fig:Energy_scale}, \ref{fig:Consumer_Staples_scale} and
\ref{fig:Information_Technology_scale} compare the estimates obtained
from Hill's scale estimator with those from the $p K^\alpha$ method.

\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}

  \end{minipage}
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}

  \end{minipage}
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  % \begin{subfigure}[b]{0.32\linewidth}
  %   \includegraphics[width=\textwidth]
  %   {Energy_scale.pdf}
  %   \caption{``Energy''}
  %   \label{fig:Energy_scale}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.32\linewidth}
  %   \includegraphics[width=\textwidth]
  %   {Consumer_Staples_scale.pdf}
  %   \caption{``Consumer Staples''}
  %   \label{fig:Consumer_Staples_scale}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.32\linewidth}
  %   \includegraphics[width=\textwidth]
  %   {Information_Technology_scale.pdf}
  %   \caption{``IT''}
  %   \label{fig:Information_Technology_scale}
  % \end{subfigure}
  \caption{\small Scale estimates obtained from Hill's scale estimator
    and from $p K^\alpha$ . Red dots are Hill's scale estimates and
    triangles are $p K^\alpha$ estimates.
  }
\end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Energy} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Energy_alpha_K}
%% \end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Consumer Staples} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Consumer_Staples_alpha_K}
%% \end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Information Technology} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Information_Technology_alpha_K}
%% \end{figure}

% \subsection{Pickand's Estimator}
% Pickands (1975) \cite{pickands1975statistical} proposed a different
% approach to estimate the tail index than that of Hill. Assume
% $\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
% denote the quantile of $1 - 1/t$ of this distribution, i.e.
% \[
% P(X > U(t)) = {1 \over t}
% \]
% The key idea of Pickands' estimator is to observe
% \[
% \lim_{t \to \infty} {
%   U(c(t) t) - U(t)
%   \over
%   U(t) - U(t/c(t))
% } = 2^{1/\alpha}
% \]
% for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
% Pickands formulated the estimation problem in terms of the extreme
% value index $\xi$, which is the reciprocal of the tail index $\alpha$
% when the distribution of $X$ is regularly varying (cf. Embrechtsm,
% Kl\"uppelberg and Mikosch \cite{Embrechts1997}).

% Building upon the aforementioned observation and taking $c(t)$ to
% be simply constant 2, Pickands proposed the following estimator for
% $\xi$:
% \[
% \hat \xi = {1 \over \ln 2} \ln {
%   x_{(k)} - x_{(2k)}
%   \over
%   x_{(2k)} - x_{(4k)}
% }
% \]
% Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
% that Pickands' estimator was consistent in both the weak and the strong
% sense, when the sequence $k(n)$ was chosen appropriately:
% \begin{itemize}
% \item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
%   \[
%   \hat \xi \overset{P}{\to} \xi
%   \]
% \item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
%   as $n \to \infty$,
%   \[
%   \hat \xi \overset{a.s.}{\to} \xi
%   \]
% \end{itemize}
% Dekkers and de Haan \cite{dekkers1989estimation} also showed the
% asymptotic normality of the estimator, when some
% extra technical conditions were satisfied by
% $U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
% \[
% \sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
% \]
% where
% \[
% v(\xi) = {
%   \xi^2 (2^{2 \xi + 1} + 1)
%   \over
%   [2 \ln(2) (2^\xi - 1)]^2
% }
% \]
% Using Pickands' estimator, the same result of very similar tail indices
% is again obtained for the {\it Energy} and the {\it Consumer Staples}
% sectors. Figure \ref{fig:Energy_Pickands} and
% \ref{fig:Consumer_Staples_Pickands} illustrate this claim.
% \begin{figure}[htb!]
%   \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
%   \caption{Pickands' estimates of {\it Energy} stocks.}
%   \label{fig:Energy_Pickands}
% \end{figure}

% \begin{figure}[htb!]
%   \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
%   \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
%   \label{fig:Consumer_Staples_Pickands}
% \end{figure}

%% \section{Hoga Test}
%% TODO: Show the test statistics of Hoga test applied to a sequence of
%% two concatenated return series. A changes in the scale is characterized
%% by a smooth bump in the test statistics while a change in the tail
%% index is characterized by a jump. Derive a theoretical result to
%% support this?

%% What is the probability of rejecting the null hypothesis when the test
%% is applied to $(X_1, X_2)$ where $X_1$ and $X_2$ have tail indices
%% $\kappa_1 \neq \kappa_2$? The probability of wrongly rejecting the
%% null hypothesis must be lower than $\alpha$.
%% \textcolor{red}{Review the test of Hoga}

% \section{Data Analysis} \label{sec:data_analysis}
% \subsection{Tail index and scale parameter}
% In this section we apply the statistical tools reviewed in the
% previous sections to real stock return series, denoted $X_t$
% hereafter. We are interested in finding out the tail indices and the
% scale parameters of the stationary distribution of $|X_t|$.

%% \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
%% Figure \ref{fig:Energy_Hill_upper} shows the Hill estimates of the
%% tail indices of 32 stocks in the {\it Energy} sector of S\&P 500
%% index; figure \ref{fig:Consumer_Staples_Hill_upper} shows the same of
%% the {\it Consumer Staples} sector.
%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Energy_Hill_upper.pdf}
%%   \caption{Hill estimates of the upper tail indices of 32 stocks'
%%     daily return series in the {\it Energy} sector of Standard \& Poor
%%     500 index. Each circle coresponds to a Hill estimate; the grey
%%     triangles above and below it mark the 97.5\% and 2.5\% quantiles
%%     of its approximate normal distribution.
%%     The low and high red lines mark the median of the 2.5\% quantiles
%%     and median of the 97.5\% quantiles, respectively.
%%     The data are from {\it Yahoo Finance} and the labels on
%%     the horizontal axis are Yahoo symbols of the stocks. The data span
%%     from 2010-01-01 to 2015-01-01 and comprise $n=1304$
%%     observations. $k=(\ln n)^2 = 49$ upper order statistics are used
%%     for computing $\alpha_n$. 
%%   }
%%   \label{fig:Energy_Hill_upper}
%% \end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Consumer_Staples_Hill_upper.pdf}
%%   \caption{Hill estimates of the upper tail indices of 30 stocks'
%%     daily return series in the {\it Consumer Staples} sector of
%%     Standard \& Poor 500 index.
%%   }
%%   \label{fig:Consumer_Staples_Hill_upper}
%% \end{figure}

%% For our purpose, we have excluded a few stocks, though:
%% \begin{figure}[htb!]
%%   \centering
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{WBA_price.pdf}
%%     \caption{{\it Walgreens Boots Alliance} prices}
%%     \label{fig:EBA_prices}
%%   \end{subfigure}
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{WFM_price.pdf}
%%     \caption{{\it Whole Foods Market} prices}
%%     \label{fig:WFM_prices}
%%   \end{subfigure}
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{EL_price.pdf}
%%     \caption{{\it Estee Lauder Cos.} prices}    
%%     \label{fig:EL_prices}
%%   \end{subfigure}
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{CL_price.pdf}
%%     \caption{{\it Colgate-Palmolive} prices}    
%%     \label{fig:CL_prices}
%%   \end{subfigure}
%%   \caption{Plot of prices of {\it Consumer Staples} stocks
%%     excluded from Hill estimates}
%% \end{figure}

%% \begin{itemize}
%% \item {\it Walgreens Boots Alliance}:

%%   As shown in figure (\ref{fig:EBA_prices}),
%%   this stock was traded only in a small fraction of the bank
%%   days and it is hence no surprise that the price varied
%%   tremendously.
%% \item {\it Whole Foods Market}, {\it Estee Lauder Cos.} and {\it
%%   Colgate-Palmolive}

%%   As seen in figures (\ref{fig:WFM_prices}), (\ref{fig:EL_prices}) and
%%   (\ref{fig:CL_prices}),
%%   the prices of these stocks were reduced by half during the
%%   period under consideration. But the price changes are the
%%   result of emission of new shares rather than random fluctuations.
%% \end{itemize}

%% Figure \ref{fig:Energy_Hill_upper} and
%% \ref{fig:Consumer_Staples_Hill_upper}
%% are suggestive of the conclusion that excluding price
%% movements due to deterministic events, tail indices of stock return
%% series tend to have the same value, within a reasonable confidence
%% interval.

%% If we look instead at the scales, differences among the stock return
%% series are however rather large. Figure (\ref{fig:Energy_Hill_Scales})
%% and (\ref{fig:Consumer_Staples_Hill_Scales}) show the Hill estimates of
%% the scale parameters of the stocks in the S\&P 500 {\it Energy} and
%% {\it Consumer Staples} sectors, respectively.
%% \begin{figure}[htb!]
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{Energy_Hill_scales.pdf}
%%     \caption{{\it Energy} scale estimates}    
%%     \label{fig:Energy_Hill_Scales}
%%   \end{subfigure}
%%   \begin{subfigure}{0.45\textwidth}
%%     \includegraphics[width=\textwidth]{Consumer_Staples_Hill_scales.pdf}
%%     \caption{{\it Consumer Staples} scale estimates}    
%%     \label{fig:Consumer_Staples_Hill_Scales}
%%   \end{subfigure}
%%   \caption{Black boxes are Hill estimates of the scale parameters of
%%     the stock return series, while red circles are the scale estimates
%%     of simulated iid. Student's $t$ series with a degree of freedom
%%     equal to the mean value of the estimated tail indices of the stock
%%     return series. The estimates of the Student's $t$ series are
%%     included for comparison with those of the return series.}
%% \end{figure}
%% Clearly the estimated scale parameters can differ by several orders of
%% magnitude.

>>>>

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
%\usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

In Section~\ref{sec:4} we summarize the discussion of the previous
sections.

@@

For convenience, we denote the scale parameter by $A$. It is the
constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
Hill \cite{hill1975simple} put forward an estimator of $A$ using his
tail index  estimator:
\[
\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
\]
where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
is the Hill tail index estimator.
Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in figure
\ref{fig:sectors_parameters}, along with estimates obtained
using the hybrid method described in section
\ref{sec:hybrid_estimation}.

@@

Routledge and Zin \cite{routledge2010generalized} assumed
a power-law utility function of $C$:
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
increasing and concave. Before proceeding any further, it is worth
stating the following simple observation:
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x; \theta)$ parameterized by
  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
  $\E h(X) < \infty$ and
  \begin{eqnarray*}
    {\partial \over \partial \theta}\left[
      F(b) - F(a)
    \right] = 0 \\
    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
    \quad \forall x \in (a, b)
  \end{eqnarray*}
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \end{enumerate}
\end{lemma}
\begin{remark}
  \label{remark:I}
  Two other cases follow trivially from lemma \ref{lemma:I}:
  \begin{enumerate}
  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
    satisfies the
    same conditions of the 1st case of lemma \ref{lemma:I},
    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
  \item By the same argument, if $h(\cdot)$ is decreasing and
    ${\pd f \over \pd \theta}$ satisfies the same conditions of the 2nd case of
    lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
  \end{enumerate}
\end{remark}
Now for simplicity we assume the investor initially has one unit of
wealth. Let $r > 0$ be the return on the risk free state bond and $X$
be the return on equity. Portfolio weights are $\left(1-\phi\right)$
and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  C(X) = (1 - \phi) e^r + \phi e^X
\end{equation}
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q = q(r, \phi) &:=& \ln \left( {
      \delta v - (1 - \phi) e^r
      \over
      \phi
    } \right) = \ln\left(
    e^r + {\delta v - e^r \over \phi}
  \right)
\end{eqnarray*}
Note $C < \delta v$ implies $X < q$. Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher return from the risky asset than he is guarranteed from the
riskless bond. In our notation, this means $\delta v > e^r$, which
implies $q(r, \phi) > r > 0$. The right-hand-side of
\eqref{eq:functional} is a functional of $\phi$ and of the distribution
function of $X$ via \eqref{eq:xxie1}. For convenience we denote it
$\mathcal U(\cdot, \cdot)$, i.e.
\begin{equation}
  \label{eq:U_functional}
  \mathcal U(F, \phi)
  = 
  \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\end{equation}
In the rest of this section, we use $F(\cdot)$ to denote the
distribution function of $X$ and use $f(\cdot)$ to denote its density
function. Let
\begin{equation}
  \label{eq:G_functional}
  G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)  
\end{equation}

@@

\subsection{Hybrid fitting of the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:X_distr}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:X_distr}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$. Regularity
conditions on the maximum likelihood estimator are easily checked for
the distribution \eqref{eq:X_distr}.
The density function of $X$ conditional on $X < 0$ is
\[
f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
\]
Clearly, the log-likelihood function of the $K$ parameter is
\[
l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
\]
where, as is just mentioned, we pre-estimate $\alpha$ with Hill
estimator and regard it as known at this step. With $n$ observations
$X_1, \dots, X_n$, the log-likelihood function of $K$ is
\[
L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
\]
To obtain the maximum likelihood estimation, we need to solve the
equation ${\pd L \over \pd K} = 0$, which expands to
\[
{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
\]
This equation has to be solved numerically.
The Fisher information of $K$ follows as
\begin{eqnarray*}
  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
  &=& \alpha K^{\alpha - 2}
  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
  &=& {\alpha \over (2 + \alpha) K^2}
\end{eqnarray*}
So the maximum likelihood estimator, call it $\hat K$, is
asymptotically normal:
\[
\sqrt n (\hat K - K) \to N \left(
  0, \left( 1 + {2 \over \alpha} \right) K^2
\right)
\]

@@

\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_p.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
  \label{fig:sectors_parameters}
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
Comparing this expression with \eqref{eq:lower_tail_expansion}, one
can immediately identity $p K^\alpha$ as the scale parameter. Figure
\ref{fig:scale_estimates} compare the estimates obtained
from Hill's scale estimator with those from the $p K^\alpha$ method.

\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}
  \end{minipage}
  \begin{minipage}{0.32\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  \caption{\small Scale estimates obtained from Hill's scale estimator
    and from $p K^\alpha$ . Red dots are Hill's scale estimates and
    triangles are $p K^\alpha$ estimates.
  }
  \label{fig:scale_estimates}
\end{figure}
